---
title: "Prob2"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Part A
We start by reading and visualizing the data. We are in a contest of 
paired data (same CDs) and we want to test if the prices and conditions of the CDs are different. 
```{r cars, echo=FALSE}

disco = read.table("discomaniac.txt", header = T)
lipstick = read.table("lipsticks.txt", header = T)


diff = disco
diff$price = disco$price - lipstick$price
diff$media.condition = disco$media.condition - lipstick$media.condition
plot(diff$price, diff$condition, main = "Differences in prices and conditions", xlab = "Prices", ylab = "Conditions")
diff = diff[, c("price", "media.condition")]
```
We will perform the test and then at point b) we will verify assumptions. 

```{r, echo=TRUE}
x.mean = colMeans(diff)
S = cov(diff)
n = 20
p = 2
alpha = 0.05
cfr.fisher = (n-1)*p/(n-p)*qf(1-alpha, p, n-p)
T2 = n*(x.mean %*% solve(S) %*% x.mean)
T2>cfr.fisher
p_value = 1 - pf(T2*(n-p)/p/(n-1), p, n-p)
p_value
```
The null hypothesis is rejected. The prices and conditions of the CDs are different. We will need to see the ellipse (or CI) for commenting this data. 

## Part B
We need to test the multivariate normality of the dataset of the differeces.
We can see that the assumption is met. 
```{r, echo=TRUE}

library(MVN)
mvn(diff)$multivariateNormality
# The assumption of multivariate normality is met. 

```

## Part C
We will plot the data and the confidence ellipse. As we can see the point c(0, 0) is not contained in the ellipse (it is so far that is not possible to see it in the graph neither)
```{r, echo=TRUE}

# c)
library(car)
plot(diff)
ellipse(center = x.mean, shape = S/n, radius = sqrt(cfr.fisher), col = "red")
abline(h = 0, lty = 3, col=  "gray")
abline(v = 0, lty = 3, col=  "gray")

# d) 

```

## Part D
Bonferroni simultaneous confidence intervals for the means.
```{r, echo=TRUE}
alpha = 0.05
k = 4

BCI_means = cbind(
  inf = x.mean - qt(1-alpha/2/k, n-1)*sqrt(diag(S)/n),
  center = x.mean,
  sup = x.mean + qt(1-alpha/2/k, n-1)*sqrt(diag(S)/n)
)
BCI_means

BCI_var = cbind(
  inf = (n - 1) * diag(S) / qchisq(1 - alpha/(2*k), n - 1),
  center = diag(S) ,
  sup = (n - 1) * diag(S) / qchisq(alpha/(2*k), n - 1)
)  
BCI_var



```
We can conclude that there is no evidence to state that one shop has better
cds than the other, however, the prices are different! Shop discomaniac is more expensive. I would go to shop lipstick. (since theere is no prof that it has worst conditions. )
```{r}