---
title: "Exe3"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exe3
We start by reading and visualizing the data:

```{r, echo=FALSE}
tatu = read.table("tattoo.txt", header = TRUE)

plot(tatu[,1:3], col = as.numeric(factor(tatu$method)))
tatu$method = factor(tatu$method)
```
In red we can see units associated to handmade tattoo, in black to machine. 
We can see that price seems correlated with dimension, and ncolors.
We start by fitting the linear model:
```{r, echo=TRUE}
tatu$method = factor(tatu$method)
m0 = lm(price ~ .*method, data = tatu)
summary(m0)
coef(m0)
sigma = summary(m0)$sigma
```
We will know see if hypotesis are satisifed. We need: 
- normality of residuals (for intervals)
- independece of residuals and homoschedasiticty
```{r, echo=FALSE}
par(mfrow = c(2, 2))
plot(m0)
par(mfrow = c(1, 1))
```
all hypotesis of lm verified, also normality: we can do confidence intervals.
## Point C
We tested if we can take 
```{r, echo=FALSE}
# point c) 
# yes for the t test. 
# maybe we can remove the intercation along with methodmachine
library(car)
C
C = matrix(0, nrow = 2, ncol = 6)
C[1, 4] =  C[2, 6] = 1
linearHypothesis(m0, C)
# we can remove them. 
#does ncolor have impact?
C
C = matrix(0, nrow = 3, ncol = 6)
C[1, 4] =  C[2, 6] = C[3, 3] = 1
linearHypothesis(m0, C)

# yes!
```
## Point D)
we will know fit a reduced model. Thanks to the test done at point C we can keep in the model dimension, ncolors and dimension:method variables. 
```{r, echo=TRUE}
# point d)
m1 = lm(price ~ dimension + ncolors + dimension:method, data = tatu)
summary(m1)
coef(m1)
summary(m1)$sigma
```
## Point E)
The fixed cost to pay every time you want a tatto is in the interval [116, 121]. See below for details. 
```{r, echo=FALSE}
# point e)
Z0.new = data.frame(dimension=0, ncolors=0, method = "machine")
predict(m1, Z0.new, interval = "confidence", level = 1-0.05/2)

Z0.new = data.frame(dimension=6.5, ncolors=1, method = "handmade")
predict(m1, Z0.new, interval = "confidence", level = 1-0.05/2)
```
