---
title: "Problem1"
output: html_document
date: "2024-06-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We start by reading data and visualizing it.
The transformation I am proposing consists in transforming the time in velocity: in this way the higher the value the better the performance.
```{r echo=FALSE}
data = read.table("decathlon.txt", header = T)
head(data)
plot(data[,3:12], col = as.numeric(factor(data$country)), pch = 16)
# i will transform the time in velocity: in this way the higher the value the better the performance
data$X100m = 100/data$X100m
data$X400m = 400/data$X400m
data$X110m_hurdles = 110/data$X110m_hurdles
data$X1500m = 1500/data$X1500m
```

We will now plot the data and interpret the results. I propose to keep the first 3 principal components
- PC1: is a weighted average of "running" and jumping variables: all velocity variables, high jump and long jump. 
- PC2: is a weighted average with negative weights of results in throwing events: shot put, discus throw, javelin throw. 
- Probably by looking at this first two PCs we are able to distinguish between sprinters and throwers.
- PC3: its seems a contrast between pole vault and highjump vs X100m (postive value). 


```{r, echo=FALSE}
# we will now plot the data
boxplot(scale(data[,3:12], center = TRUE, scale = FALSE))
# we see that we need to standardize the data, since shot put, discus throw and javelin throw 
# have a much higher variance than the other variables. 

fit.pca = princomp(scale(data[,3:12]))
summary(fit.pca)
# we can see that we need at least 3 PCs to explain 70% of the variance.4 if we want to reach the 80% threshold 

library(MiniR)
pc.interpretation(fit.pca, data = data[,3:12])

```
#Screeplot
From the screeplot we are able to distinguish between sprinters(high values of PC1 scores) and throwers (negative value of PC2).
 we repeat the biplot with the plot of second and third PC
```{r, echo=FALSE}

# point c)
biplot(fit.pca)
# f

```
#point d)
see point a) for screeplot. 
```{r, echo=FALSE}

# point d)
# see point a  for screeplot. 
# I will keep just the first three pcs, we could keep two, but we may lose informaion 
# on pale vault and high jump.

```

#point e)
It seem a pretty balanced athlete, with a good performance in all the events but the throwing ones.

```{r, echo=FALSE}


Z0.new = data.frame(X100m = 100/10.81, long_jump = 7.46, shot_put = 14.56, high_jump = 2.02, X400m = 400/47.90, X110m_hurdles = 110/14.44, discus_throw = 43.04, pole_vault = 4.90, javelin_throw = 57.24, X1500m = 1500/281.63)
Z0.new = (Z0.new - colMeans(data[,3:12]))/sqrt(diag(var(data[,3:12])))
Z0.new = as.matrix(Z0.new)
Z0.new = Z0.new %*% fit.pca$loadings[,1:3]
print("new athlete scores: ")
print(Z0.new)
biplot(fit.pca)
points(Z0.new[,1], Z0.new[,2], col = "blue", pch = 2, cex = 2)
# It seem a pretty balanced athlete, with a good performance in all the events but the throwing ones.



```
